Real Time Fluid Simulation \\ \smaller using Smoothed-Particle Hydrodynamics and OpenGL

### 2014-12-14 paper/paper.tex
added cubic spline implementation and derivative

>>>>

As a kernel function we used the standard kernel, the \emph{cubic spline}
\begin{align*}
	W_3(r,h) := \frac{1}{\pi h^3}
	\begin{cases}
		1 - \frac{3}{2} \xi^2 + \frac{3}{4} \xi^3, &\text{for } 0 \le \xi < 1 \\
		\frac{1}{4} (2-\xi)^3, &\text{for } 1 \le \xi < 2 \\
		0, &\text{otherwise}
	\end{cases}
\end{align*}
%
where $\xi := \frac{r}{h}$. It can be seen clearly that the support of this function is $2h$. \\
The derivative is
%
\begin{align*}
	\nabla W_3(r,h) &= \frac{\partial}{\partial r} W_3(r,h) + \frac{\partial}{\partial h} W_3(r,h)
\end{align*}
%
\begin{itemize}
	\item for $0 \le \xi < 1$
		\begin{align*}
			\frac{\partial}{\partial r} W_3(r,h)|_{0 \le \xi < 1}
			&= \frac{1}{\pi h^3} \frac{1}{h} \left( 3 \xi + \frac{9}{4} \xi^2 \right)
		\end{align*}
		%
		\begin{align*}
			\frac{\partial}{\partial h} W_3(r,h)|_{0 \le \xi < 1}
			&= \frac{-3}{\pi h^4} \left( 1 - \frac{3}{2} \xi^2 + \frac{3}{4} \xi^3 \right) \\
			&\quad + \frac{1}{\pi h^3} \frac{1}{r} \left( (-2) \frac{3}{2} \xi^3 + (-3)\frac{3}{4} \xi^4 \right) \\
			&= - \frac{3}{\pi h^4} \left( 1 - \frac{3}{2} \xi^2 + \frac{3}{4} \xi^3 \right) \\
			&\quad + \frac{1}{\pi h^3} \frac{1}{r} \left( -3 \xi^3 -\frac{9}{4} \xi^4 \right)
		\end{align*}
		%
		\begin{align*}
			\Rightarrow \nabla W_3(r,h)|_{0 \le \xi < 1}
		&= \frac{1}{\pi h^4} \left( 3 \xi + \frac{9}{4} \xi^2 \right) \\
			&- \frac{3}{\pi h^4} \left( 1 - \frac{3}{2} \xi^2 + \frac{3}{4} \xi^3 \right) \\
			&\quad + \frac{1}{\pi h^3} \frac{1}{r} \left( -3 \xi^3 -\frac{9}{4} \xi^4 \right) \\
		&= \frac{3}{\pi h^4} \left( -1 + \xi + \frac{9}{4} \xi^2
			- \frac{3}{4} \xi^3 \right) \\
			&\quad - \frac{1}{\pi r h^3} \xi^3 \left( 3 + \frac{9}{4} \xi \right)
	\end{align*}
	\item for $1 \le \xi < 2$
		\begin{align*}
			\frac{\partial}{\partial r} W_3(r,h)|_{1 \le \xi < 2}
			&= \frac{1}{4} 3(2-\xi)^2 \cdot \left( -\frac{1}{h} \right) \\
			&= -\frac{3}{4h} (2-\xi)^2
		\end{align*}
		%
		\begin{align*}
			\frac{\partial}{\partial h} W_3(r,h)|_{1 \le \xi < 2}
			&= \frac{1}{4} 3(2-\xi)^2 \cdot \left( +\frac{\xi}{h} \right) \\
			&= \frac{3 \xi}{4 h} (2-\xi)^2
		\end{align*}
		\begin{align*}
			\Rightarrow \nabla W_3(r,h)|_{1 \le \xi < 2}
			&= -\frac{3}{4h} (2-\xi)^2
			+ \frac{3 \xi}{4 h} (2-\xi)^2 \\
			&= \frac{3}{4h} (2-\xi)^2 (\xi - 1)
		\end{align*}
	\item otherwise
		\begin{align*}
			\nabla W_3(r,h)|_{\text{otherwise}} = 0
		\end{align*}
\end{itemize}
%
The following two code snippets show how these formulas are implemented in our simulation:
\begin{lstlisting}[caption=Cubic Spline]
// Cubic spline function
inline float W3(float r, float h) {
	float W = 0;
	float zeta = r/h;
	if(0 <= zeta && zeta <= 2) {
		W = (zeta < 1 ? 
				1 - .75*zeta*zeta*(2+zeta) :
				.25*(2-zeta)*(2-zeta)*(2-zeta)
		);
	}
	W /= (M_PI*h*h*h);
	return W;
}
\end{lstlisting}
%
\begin{lstlisting}[caption=Derivative of Cubic Spline]
Hello World
\end{lstlisting}
